---
title: "Australia Rain Fall Prediction"
author: "Arup Ghosh"
date: "3/6/2021"
output: html_document
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r libraries, include=TRUE}
library("bnlearn")
#library("ggdag")
library("tidyverse")
# library("visNetwork")
library(lubridate)
```
# Application domain: Rain Forecast

Rain Forecast depends on:

- Maximum Temperature $MT$.

- Sunshine $SS$.

- Humidity at 9AM $H9AM$. Directly influenced by Maximum Temperature ($MT$) and Sunshine ($SS$).

- Humidity at 3PM $H3PM$. Directly influenced by Sunshine ($SS$).

- Rain in Tomorrow $RT$. Directly influenced by Humidity at 9AM ($H9AM$) and Humidity at 3PM $H3PM$.

- Create the graphical model using the `model2network()` method.

## Graphical model

```{r, echo=TRUE}
dag.bnlearn <- model2network("[MT][SS][H9AM|MT:SS][H3PM|SS][RT|H9AM:H3PM]")
dag.bnlearn
```
## Graphical model Visualization
```{r, echo=TRUE}
#plot(dag.bnlearn)
```

## Distributions of variables

- $\mathrm{MT} \sim N \left(24, \,\, 7^{2}\right)$

- $\mathrm{SS} \sim N \left(9, \,\, 4^{2}\right)$

- $\mathrm{H9AM} \mid \mathrm{MT}, \mathrm{SS} \sim N \left( 110 - 1.37 \cdot \mathrm{MT} - 2.13 \cdot \mathrm{SS}, \,\, 15^{2}\right)$

- $\mathrm{H3PM} \mid \mathrm{SS} \sim N \left(73.47 - 4.42 \cdot \mathrm{SS}, \,\, 13.31^{2}\right)$

- $\mathrm{RT} \mid \mathrm{H9AM}, \mathrm{H3PM} \sim N\left(-0.08 -0.0042\cdot  \mathrm{~H9AM} + 0.123 \cdot \mathrm{~H3PM}, \,\,  0.33^{2}\right)$

&nbsp; &nbsp; &nbsp;

## Exploring independencies using the `dsep` function of the `bnlearn` library


```{r, echo=TRUE}
nodes <- nodes(dag.bnlearn)
  for (n1 in nodes) {
    for (n2 in nodes) {
      if (dsep(dag.bnlearn, n1, n2))
        cat(n1, "and", n2, "are independent.\n")
    }
  }
```

- Independence is symmetric.

```{r, echo=TRUE}
dsep(dag.bnlearn, "MT", "MT")
dsep(dag.bnlearn, "SS", "SS")
dsep(dag.bnlearn, "H9AM", "H9AM")
```

- Find which pairs of variables are conditionally independent given $H9AM$

```{r, echo=TRUE}
nodes <- nodes(dag.bnlearn)
for (n1 in nodes[nodes != "H9AM"]) {
  for (n2 in nodes[nodes != "H9AM"]) {
    if (n1 < n2) {
      if (dsep(dag.bnlearn, n1, n2, "H9AM"))
        cat(n1, "and", n2, "are independent given H9AM.\n")
    }
  }
}
```



- Find which pairs of variables are conditionally independent given $H3PM$

```{r, echo=TRUE}
nodes <- nodes(dag.bnlearn)
for (n1 in nodes[nodes != "H3PM"]) {
  for (n2 in nodes[nodes != "H3PM"]) {
    if (n1 < n2) {
      if (dsep(dag.bnlearn, n1, n2, "H3PM"))
        cat(n1, "and", n2, "are independent given H3PM.\n")
    }
  }
}
```

- Is there a path from $MT$ to $RT$?

```{r, echo=TRUE}
bnlearn::path(dag.bnlearn, from = "MT", to = "RT")
```

- Is there a path from $SS$ to $RT$?

```{r, echo=TRUE}
bnlearn::path(dag.bnlearn, from = "SS", to = "RT")
```
# Specifying the joint probability distribution
```{r, echo=TRUE}
distMT <- list(coef = c("(Intercept)" = 24), sd = 7)
distSS <- list(coef = c("(Intercept)" = 9), sd = 4)
distH9AM <- list(coef = c("(Intercept)" = 110, MT = -1.37, SS = -2.13), sd = 15)
distH3PM <- list(coef = c("(Intercept)" = 73.47, SS = -4.42), sd = 13.31)
distRT <- list(coef = c("(Intercept)" = 0, H9AM = 0, H3PM = 0), sd = 0.33)
dist.list = list(MT = distMT, SS = distSS, H9AM = distH9AM, H3PM = distH3PM, RT = distRT)
```
# Gaussian Bayesian networks (GBNs)

- All variables are normally distributed.

- Root nodes (i.e., nodes without any parent), are characterized by their marginal distributions.

- The conditioning effect of the parent nodes is given by an additive linear term in the mean; variance is not affected -- each node has a variance that is specific to that node and does not depend on the values of the parents.

- A node's local distribution expressed as a Gaussian linear model (an intercept and the node's parents as
explanatory variables; no interaction terms).

```{r, echo=TRUE}
gbn.bnlearn <- custom.fit(dag.bnlearn, dist = dist.list)
```


```{r, echo=TRUE}
gbn.bnlearn$SS
```


```{r, echo=TRUE}
gbn.bnlearn$RT
```
## rbmn R package for GBNs

- Convert the `gbn.bnlearn` (a `bn.fit` object) to `rbmn` native format.

```{r, echo=TRUE}
library(rbmn)
gbn.rbmn <- bnfit2nbn(gbn.bnlearn)
```
```{r, echo=TRUE}
gema.rbmn <- nbn2gema(gbn.rbmn)
mn.rbmn <- gema2mn(gema.rbmn)
print8mn(mn.rbmn)
```


- Assumption: The joint distribution of all nodes (i.e., the global distribution) is a **multivariate normal distribution**.

\begin{equation}
f(\mathrm{MT}, \mathrm{SS}, \mathrm{H9AM}, \mathrm{H3PM}, \mathrm{RT}) = f(\mathrm{MT}) \, f(\mathrm{SS})  \, f(\mathrm{H9AM} \mid \mathrm{MT}, \mathrm{SS}) \,  f(\mathrm{H3PM} \mid \mathrm{SS}) \, f(\mathrm{RT} \mid \mathrm{H9AM}, \mathrm{H3PM})
\end{equation}

- Numeric derivation of the parameters of the **multivariate normal distribution**. 

```{r, echo=TRUE}
gema.rbmn <- nbn2gema(gbn.rbmn)
mn.rbmn <- gema2mn(gema.rbmn)
print8mn(mn.rbmn)
```
- Structure of the `mn.rbmn` object

```{r, echo=TRUE}
str(mn.rbmn)
```





# Read the dataset

```{r, include=TRUE}
rain.df <- read.csv("Data/RainFallAus/weatherAUS.csv",nrows = 8000)[ ,c("MaxTemp", "Sunshine","Humidity9am","Humidity3pm", "RainTomorrow" )]

```

# Training Data and Test Data split
```{r, echo=TRUE}
dt = sort(sample(nrow(rain.df), nrow(rain.df)*.8))
train<-rain.df[dt,]
test<-rain.df[-dt,]
```


```{r, echo=TRUE}
dim(train)
str(train)
```
```{r, echo=TRUE}
library(GGally)
## Max Temperature (MT)
#mean
mean(train$MaxTemp,na.rm = TRUE)
#standard deviation
sd(train$MaxTemp,na.rm = TRUE)

## Sunshine (SS)
#mean
mean(train$Sunshine,na.rm = TRUE)
#standard deviation
sd(train$Sunshine,na.rm = TRUE)


## Humidity 9AM (Humidity9am)
model_h9am <- lm(train$Humidity9am ~ train$MaxTemp + train$Sunshine , na.rm = TRUE)
#mean
coef(model_h9am)
#standard deviation
summary(model_h9am)$sigma

## Humidity 3PM (Humidity3pm)
model_h3pm <- lm(train$Humidity3pm ~ train$Sunshine , na.rm = TRUE)
#mean
coef(model_h3pm)
#standard deviation
summary(model_h3pm)$sigma

## Rain Tomorrow (RT)
model_rt <- lm(train$RainTomorrow ~ train$Humidity9am + train$Humidity3pm , na.rm = TRUE)
#mean
coef(model_rt)
#standard deviation
summary(model_rt)$sigma

```
- For continuous data, `bn.fit` implements only **mle**  estimator.


```{r, echo=TRUE}
str(train)

train$Sunshine <- as.numeric(train$Sunshine, na.rm = TRUE)
train$Humidity9am <- as.numeric(train$Humidity9am, na.rm = TRUE)
train$Humidity3pm <- as.numeric(train$Humidity3pm, na.rm = TRUE)
train$RainTomorrow <- as.numeric(train$RainTomorrow, na.rm = TRUE)
#train[] <- lapply( train, factor)
str(train)

names(train)
names(train)[names(train) == "MaxTemp"] <- "MT"
names(train)[names(train) == "Sunshine"] <- "SS"
names(train)[names(train) == "Humidity9am"] <- "H9AM"
names(train)[names(train) == "Humidity3pm"] <- "H3PM"
names(train)[names(train) == "RainTomorrow"] <- "RT"

```
# Estimating the model parameters: Correlation coefficients

- We assume that the structure of the GBN is known.

- 200 observations are generated from the GBN and saved in a data frame (cropdata1).


```{r, echo=TRUE}
set.seed(4567)
raindata1 <- rbn(gbn.bnlearn, n = 200)
set.seed(1234)
raindata2 <- rbn(gbn.bnlearn, n = 20000)
str(raindata2)
```
```{r, echo=TRUE}
est.para <- bn.fit(dag.bnlearn, method = "mle", data = raindata1,na.rm = TRUE)
```
```{r, echo=TRUE}
est.para$RT <- lm(RT ~ H9AM + H3PM, data = raindata1, na.rm = TRUE)
est.para$H9AM <- lm(H9AM ~ MT + SS, data = raindata1, na.rm = TRUE)

```
- Ridge regression for the random variable $RT$

```{r, echo=TRUE}
library(penalized)
est.para$RT <- penalized(RT ~ H9AM + H3PM, lambda1 = 0, lambda2 = 1.5, data = raindata1)
```
```{r, echo=TRUE}
est.para$RT
```
```{r, echo=TRUE}
est.para$RT <- lm(RT ~ H9AM + H3PM - 1, data = raindata1)
est.para$RT
```
- Parameter estimates are based only on the subset of the original data frame spanning the considered node and its parents.

```{r, echo=TRUE}
lmRT <- lm(RT ~ H9AM + H3PM, data = raindata1[, c("H9AM", "H3PM", "RT")])
coef(lmRT)
```
```{r, echo=TRUE}
confint(lmRT)
```
```{r, echo=TRUE}
cormat <- cor(raindata1[, c("RT", "H3PM", "H9AM")])
cormat
```
```{r, echo=TRUE}
library(corpcor)
invcor <- cor2pcor(cormat)
dimnames(invcor) <- dimnames(cormat)
invcor
```
## Structure learning



```{r, echo=TRUE}
stru1 <- iamb(raindata1, test = "cor")
stru1
```
```{r, echo=TRUE}
library(DiagrammeR)
DiagrammeR::mermaid("
  graph TB
  mt((MT)) --> h9am((H9AM))
  ss((SS)) --> h9am((H9AM))
  ss((SS)) --> h3pm((H3PM))
  h9am((H9AM)) --> rt((RT))
  h3pm((H3PM)) --> rt((RT))
")
```
## Network scores
- Bayesian Gaussian equivalent score (BGe) -- posterior probability score in common use is an uniform prior over the space of DAGs and of the parameters.


```{r, echo=TRUE}
score(dag.bnlearn, data = raindata2, type = "bic-g")
```

```{r, echo=TRUE}
score(dag.bnlearn, data = raindata2, type = "bge")
```
```{r, echo=TRUE}
print8nbn(gbn.rbmn)
```


```{r, echo=TRUE}
print8gema(gema.rbmn)
```