---
title: "CreditCard_Predict_v1"
author: "Arup Ghosh"
date: "2/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


# A Probabilistic Graphical Model (PGM)
```{r libraries, include=FALSE}
library(ggplot2)
library(dplyr)
library(reshape2)
library(gridExtra)
library(GGally) # ggpairs() for scatterplot matrix
library(RColorBrewer)
library("tidyverse") # a better data frame
library("visNetwork") # for constructing BN graph using a Javascript library
library("bnlearn") # for Bayesian network learning
```

# Problem statement: Who gets Platinum Credit Card?

Demographic predictors

  - Age (A):  **ADULT** for individuals between 30 and 60 years old; and **MATURE** for people older than 60.
  
Socioeconomic indicators

  - Education level (E): the highest level of education completed by the individual: Doctorate,Post Graduate,Graduate (**HIGH**), High School,uneducated (**LOW**).
  - Income Category (I): more than 60K (**HIGH**) or less than 60K (**LOW**) 
  - Marital Status (M):  **MARRIED** or **SINGLE**

Prediction target

  - Credit Card type (C):  **SILVER**, **GOLD**, or **PLATINUM**

## Read the dataset-Data Aalysis
```{r, include=TRUE}
credit_card.df <- read.csv("Data/credit_card/BankChurners.csv")
class(credit_card.df)
names(credit_card.df)
dim(credit_card.df)
str(credit_card.df)
summary(credit_card.df)
```

# Boxplots
```{r, echo=FALSE, message=FALSE, warning=FALSE}

credit_card.df$E = as.factor(credit_card.df$I)

ggplot(aes(x = E, y = I), data = credit_card.df) +
  geom_boxplot(color="aquamarine4") +
  geom_jitter(alpha = 0.1, color="firebrick")+
  coord_cartesian(ylim=c(0, 35))

ggplot(aes(x = A, y = I), data = credit_card.df) +
  geom_boxplot(color="aquamarine4") +
  geom_jitter(alpha = 0.1, color="firebrick")+
  coord_cartesian(ylim=c(0, 35))

ggplot(aes(x = A, y = M), data = credit_card.df) +
  geom_boxplot(color="aquamarine4") +
  geom_jitter(alpha = 0.1, color="firebrick")+
  coord_cartesian(ylim=c(0, 35))

ggplot(aes(x = M, y = C), data = credit_card.df) +
  geom_boxplot(color="aquamarine4") +
  geom_jitter(alpha = 0.1, color="firebrick")+
  coord_cartesian(ylim=c(0, 35))

ggplot(aes(x = I, y = C), data = credit_card.df) +
  geom_boxplot(color="aquamarine4") +
  geom_jitter(alpha = 0.1, color="firebrick")+
  coord_cartesian(ylim=c(0, 35))

```

## Define the Bayesian network

```{r define-nodes, echo=TRUE}
dag <- empty.graph(nodes = c("A", "E", "I", "M", "C"))
```

```{r, echo=TRUE}
dag
```

```{r define-edges, echo=TRUE}
dag <- set.arc(dag, from = "E", to = "I")
dag <- set.arc(dag, from = "A", to = "I")
dag <- set.arc(dag, from = "A", to = "M")
dag <- set.arc(dag, from = "I", to = "C")
dag <- set.arc(dag, from = "M", to = "C")
```

```{r, echo=TRUE}
dag
```
# Exploring the bn object

```{r, echo=TRUE}
nodes(dag)
```

```{r, echo=TRUE}
arcs(dag)
```


# Joint probability as a product of conditional probabilitiies

Use the `modelstring` function to generate the product of conditional probabilities.

```{r, echo=TRUE}
modelstring(dag)
```
# Specify joint probability distribution over the random variables


```{r, echo=TRUE}
A.lv <- c("ADULT", "MATURE") # age
E.lv <- c("HIGH", "LOW") # education category
I.lv <- c("HIGH", "LOW") # income category
M.lv <- c("Married", "Single") # marital status
C.lv <- c("Silver", "Gold","Platinum") # card type
```

#Marginal probability distributions for Education variables
```{r, echo=TRUE}                
E.prob <- array(c(.40,.60),dim=2,dimnames = list(E = E.lv))
E.prob
```

#Marginal probability distributions for Age variables
```{r, echo=TRUE}                
A.prob <- array(c(.40,.60),dim=2,dimnames = list(A = A.lv))
A.prob
```

#Conditional Probability Tables (CPTs) for Income Category variable are modeled using three-dimensional tables:
```{r, echo=TRUE}
I.prob <- array(c(.35,.65,.37,.63,.2,.8,.15,.85), dim = c(2, 2, 2),
            dimnames = list(I = I.lv, E = E.lv, A = A.lv))
I.prob
```

#Conditional probability distributions for Marital Status variables:
```{r, echo=TRUE}
M.prob <- array(c(.46,.54,.45,.55), dim = c(2, 2),
            dimnames = list(M = M.lv, A = A.lv))
M.prob
```

#Conditional Probability Tables (CPTs) for Credit Card Type variable are modeled using three-dimensional tables:
```{r, echo=TRUE}
C.prob <- array(c(.98,.01,.01,.98,.01,.01,.97,.02,.01,.98,.01,.01), dim = c(3, 2, 2),
            dimnames = list(C = C.lv, I = I.lv, M = M.lv))
C.prob
```
# bn.fit class
# We can combine our DAG defined earlier with CPTs to create an object of the bn.fit class:
```{r, echo=TRUE}
cpt <- list(A = A.prob, E = E.prob, I = I.prob, M = M.prob, 
        C = C.prob)

bn <- custom.fit(dag, cpt)
```
#Number of parameters of the model:
```{r, echo=TRUE}
nparams(bn)
```
#Number of edges in the model:
```{r, echo=TRUE}
arcs(bn)
```
#Print all CPTs in the model:
```{r, echo=TRUE}
bn

```


# Estimating the model parameters from an observed sample

```{r, echo=TRUE}
survey <- read.table("data/credit_card/BankChurners.csv", header = TRUE,sep=",")
```


```{r, echo=TRUE}
class(survey)
```
```{r, echo=TRUE}
str(survey)
```

```{r, echo=TRUE}
names(survey)
```

```{r, echo=TRUE}
class(survey$A)
```
```{r, echo=TRUE}
survey[] <- lapply( survey, factor)
```

```{r, echo=TRUE}
str(survey)
```


```{r, echo=TRUE}
class(survey$A)
```

```{r, echo=TRUE}
head(survey)
```

# Training Data and Test Data split
```{r, echo=TRUE}
dt = sort(sample(nrow(survey), nrow(survey)*.7))
train<-survey[dt,]
test<-survey[-dt,]
```

#In **bnlearn**, we compute MLE using the **bn.fit** function.
```{r, echo=TRUE}
options(digits = 3)
bn.mle <- bn.fit(dag, data = train, method = "mle")

bn.bayes <- bn.fit(dag, data = train, method = "bayes", 
              iss = 10)
bn.bayes$M

bn.bayes <- bn.fit(dag, data = train, method = "bayes", 
              iss = 30)
bn.bayes$M
```
#Network scores
```{r, echo=TRUE}
set.seed(456)
options(digits = 6)
score(dag, data = train, type = "bic")

score(dag, data = test, type = "bic")

rnd <- random.graph(nodes = c("E", "A", "I", "M", "C"))
modelstring(rnd)

score(rnd, data = train, type = "bic")
score(rnd, data = test, type = "bic")
```

```{r, echo=TRUE}
learned <- hc(train)
modelstring(learned)
learned2 <- hc(train, score = "bde")
options(digits=3)
arc.strength(learned, data = train, criterion = "bic")
arc.strength(dag, data = train, criterion = "bic")
```
#Plot the model
```{r, echo=TRUE}
graphviz.plot(dag)
```
